### 2021-04-18 model_test_script.Rmd

---
title: "Untitled"
author: "Hong Shi"
date: "4/17/2021"
output: pdf_document
---
```{r}
library(tidyverse)
library(here)
library(finalfit)
library(kableExtra)
library(rdrobust)
library(broom)
#install.packages("rdpower")
library(rdpower)
```

```{r}
clean_cut <- read_csv(here("inputs/data/adjust_bandwidth_data.csv"))
```

```{r}
# Histogram of the distribution of GPA from cutoff

ggplot(clean_for_summary, aes(x = dist_from_cut))+
  geom_histogram(aes(y  = stat(count)), binwidth = 0.125, color = "black", fill = "#33CCCC")+
  geom_vline(xintercept = 0, linetype = "dotted")+
  theme_minimal()+
  theme(plot.title = element_text(hjust = 0.5)) +
  ylab("Frequency Headcount")+
  xlab("Year 1 GPA minus Probation Cutoff")+
  ggtitle("Distribution of Year 1 GPA relative to Probation Cutoff for All Students")
```

```{r}
# plot to show the probation assignment
ggplot(clean_cut,aes(x = dist_from_cut, y = status, group = status, color = status))+
  geom_point(alpha = 0.1)+
  geom_smooth(method = "lm")+
  theme_minimal()+
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5))+
  labs( title = "Stuent Academic Status",
        y ="Academic Status",
        x = "Year 1 GPA minus Probation Cutoff")
```
```{r}
# Check the number of students with administrative data error reporting 
clean_for_summary %>% 
  filter((dist_from_cut < 0 & probation_year1 !=1)| (dist_from_cut >=0 & probation_year1 !=0) ) %>% 
  length()
```

```{r}
clean_cut$status <- as.factor(clean_cut$probation_year1)
ggplot(data= clean_cut, aes(x = dist_from_cut, y = GPA_year2, color = status))+
  geom_point(alpha = 0.1)+
  geom_smooth()
```
```{r}
class(clean_cut$gender)
class(clean_cut$highHS)
class(clean_cut$first_language)
clean_cut$gender <- as.factor(clean_cut$gender)
clean_cut$highHS <- as.factor(clean_cut$highHS)
clean_cut$first_language <- as.factor(clean_cut$first_language)

```
```{r}
class(clean_cut$)
class(clean_cut2$dist_from_cut)

```

```{r}
rdplot(y = clean_cut$GPA_year2, x = clean_cut$dist_from_cut, c = 0, p = 1, covs = cbind(clean_cut$gender,clean_cut$highHS, clean_cut$first_language), title = "GPA by Year 4")
```
```{r}
rdplot(y = clean_cut$left_school, x = clean_cut$dist_from_cut, c = 0, p = 1, covs = cbind(clean_cut$gender,clean_cut$highHS, clean_cut$first_language), title = "GPA by Year 4")
```

```{r}
clean_male <- clean_cut %>% 
  filter(female ==1)
class(clean_male$left_school)
clean_male$left_school <- as.factor(clean_male$left_school)
```

```{r}
test_male <- glm(left_school ~ status + dist_from_cut + highHS + first_language ,family = "binomial", data = clean_male)
summary(test_male)
```
```{r}
rdplot(y = clean_male$left_school, x = clean_male$dist_from_cut, c = 0, p = 1, covs = cbind(clean_male$highHS, clean_male$first_language), title = "GPA by Year 4")
```