### 2021-04-18 test_script.Rmd

---
title: "test script"
author: "Hong Shi"
date: "4/4/2021"
output: pdf_document
---
```{r}
library(tidyverse)
library(here)
library(haven)
library(rdrobust)
library(finalfit)
```
```{r}
?rdrobust
```

```{r}
raw_data <- read_dta(here("inputs/data/original_data.dta"))
```

```{r}
colnames(raw_data)
```
```{r}
sum(is.na(raw_data$GPA_year1))
```
```{r}
high_gpa <- 
  raw_data %>% 
  filter(GPA_year1 > 4)
```

```{r}
ggplot(raw_data, aes(x = GPA_year1))+
  geom_histogram(aes(y  = ..density..), binwidth = 0.1, color = "black", fill = "#33CCCC")+
  geom_density(aes(y = ..density..))+
  theme_minimal()
```
```{r}
ggplot(clean_data, aes(x = dist_from_cut))+
  geom_histogram(aes(y  = ..density..), binwidth = 0.02, color = "black", fill = "#33CCCC")+
  geom_density(aes(y = ..density..))+
  theme_minimal()
```

```{r}
?rdplot
```

```{r}
install.packages("rdrobust")
```
```{r}
class(raw_data$dist_from_cut)
```

```{r}
#finalfit
library(finalfit)
less_cutoff <- 
  raw_data %>% 
  filter(dist_from_cut<=0, dist_from_cut >= -0.3)
more_cutoff <- 
  raw_data %>% 
  filter(dist_from_cut > 0, dist_from_cut < 0.3 )

sample_data <- 
  raw_data %>% 
  filter(dist_from_cut >= -0.3, dist_from_cut <= 0.3)

clean_data <- 
  sample_data %>% 
  mutate(rdd_group = 
           case_when(
             (dist_from_cut >= -0.3) & (dist_from_cut < 0)~ "Academic Probation",
             (dist_from_cut >= 0) & (dist_from_cut <= 0.3)~ "In Good Standing"))

```

```{r}
test <- 
  raw_data %>% 
  filter(dist_from_cut ==0)
```

```{r}
library(VIM)
```

```{r}
?rdplot
```
```{r}
unique(clean_data$GPA_year1)
```

```{r}
summary(rdbwselect(clean_data$GPA_year1, clean_data$dist_from_cut, c=0, all = TRUE))
```
```{r}
secondyear <- clean_data %>% 
  filter(GPA_year2 != 0)
sum(is.na(secondyear$dist_from_cut))
```
```{r}
sGPA <- secondyear$GPA_year2 - secondyear$gpacutoff
head(sGPA,100)
```

```{r}
rdplot(sGPA, secondyear$dist_from_cut, c= 0, p = 3)
```
```{r}
summary(rdrobust(y = secondyear$GPA_year2, x = secondyear$dist_from_cut,
         c = 0, h = 0.3, all = TRUE, covs = as.factor(secondyear$hsgrade_pct)))
# Reference the following to add covariates
# https://stackoverflow.com/questions/57599081/fitting-a-regression-discontinuity-model-in-r-with-covariates-using-rdrobust-pac
rdrobust(y = secondyear$GPA_year2, x = secondyear$dist_from_cut,
         c = 0, h = 0.3, all = TRUE, covs = cbind(secondyear$hsgrade_pct))
test_output <- rdrobust(y = secondyear$GPA_year2, x = secondyear$dist_from_cut,
         c = 0, h = 0.3, all = TRUE, covs = cbind(secondyear$sex), scaleregul = 1)
summary(test_output)
```
```{r}
attach(secondyear)
test_output <- rdrobust(y = secondyear$GPA_year2, x = secondyear$dist_from_cut,
         c = 0, h = 0.3, all = TRUE, covs = cbind(secondyear$hsgrade_pct), scaleregul = 1)
detach(secondyear)
stargazer(test_output)
```
```{r}
stargazer(clean_data, title = "Summary Statistics")
```

```{r}
install.packages("rdd")
library(rdd)
```

```{r}
IKbandwidth(secondyear$dist_from_cut,secondyear$GPA_year2, cutpoint = 0, verbose = TRUE)
```
```{r}
test1 <- RDestimate(GPA_year2 ~ dist_from_cut, data = secondyear, cutpoint = 0, bw = 0.3 , verbose = TRUE, model = TRUE, frame = TRUE)
plot(test1)
```
```{r}
secondyear$sex <- as.factor(secondyear$sex)
class(secondyear$sex)
```

```{r}
RD_est <- function(mod, covariates) {
  RD_fit <- RDestimate(as.formula(paste(mod, covariates)), 
                       data = RD_data, cutpoint = 0)
  with(RD_fit, c(est = est[[1]], se = se[1], p = p[1]))
}
covariates <- list("No ")
```

```{r}
test1 <- RDestimate(GPA_year2 ~ dist_from_cut | sex, data = secondyear, cutpoint = 0, bw = 0.3 , verbose = TRUE, model = TRUE, frame = TRUE)
summary(test1)
```

```{r}
set.seed(853)

number_of_observation <- 1000

rdd_example_data <- tibble(person = c(1:number_of_observation),
                           grade = runif(number_of_observation, min = 78, max = 82),
                           income = rnorm(number_of_observation, 10, 1)
                           )

## We want to make income more likely to be higher if they are have a grade over 80
rdd_example_data <- 
  rdd_example_data %>% 
  mutate(income = if_else(grade > 80, income + 2, income))
```

```{r}
rdrobust(y = rdd_example_data$income, 
         x = rdd_example_data$grade, 
         c = 80, h = 2, all = TRUE) %>% 
  summary()
```
```{r}
library(lfe)
```

```{r}
groups
```

```{r}
head(clean_cut$clustervar)
```

```{r}
class(clean_cut$left_school)
clean_cut$left_school <- as.factor(clean_cut$left_school)
clean_cut
```

```{r}
# Refactor these
class(clean_cut$sex)
clean_cut$sex <- as.factor(clean_cut$sex)
clean_cut$year2_dist_from_cut
```

```{r}
lm_reg_all <- lm(year2_dist_from_cut ~ probation_year1 + dist_from_cut, data = clean_cut)
summary(lm_reg_all)
clean_cut$left_school
```
```{r}
clean_cut$probation_year1 <- as.factor(clean_cut$probation_year1)
ggplot(data= clean_cut, aes(x = dist_from_cut, y = year2_dist_from_cut, color = probation_year1))+
  geom_point(alpha = 0.1)+
  geom_smooth()
```

```{r}
lm_reg_all <- glm(left_school ~ probation_year1 + dist_from_cut , family = "binomial", data = clean_cut)
summary(lm_reg_all)
```
```{r}
test1 <- RDestimate(left_school ~ dist_from_cut , data = clean_cut, cutpoint = 0, bw = 0.3 ) 
                    #verbose = TRUE, model = TRUE, frame = TRUE)
summary(test1)
```
```{r}
clean_cut$dist_from_cut
```

```{r}
?rdrobust
test_rdd <- rdrobust(y = clean_cut$left_school, x = clean_cut$dist_from_cut ,c = 0, h = 0.3, all = TRUE)
summary(test_rdd)
```

```{r}
lm_reg <- lm(left_school ~ probation_year1 + dist_from_cut, data = subset(clean_cut, male == 1))
lm_reg1 <- lm(left_school ~ probation_year1 + dist_from_cut, data = subset(clean_cut, female == 1))
summary(lm_reg)
```

```{r}
summary(lm_reg1)
```

```{r}
class(clean_data$lowHS)
clean_data$lowHS <- as.factor(clean_data$lowHS)

```

```{r}
?sample
lowHS <- sample[clean_data$lowHS == 1, ]
highHS <- sample[clean_data$highHS == 1, ]
male <- sample[clean_data$male == 1, ]
female <- sample[clean_data$female == 1, ]
english <- sample[clean_data$english == 1, ]
noenglish <- sample[clean_data$noenglish == 1, ]
```
```{r}
table3col1 <- lm(probation_year1 ~ gpalscutoff + gpaXgpalscutoff + gpaXgpagrcutoff, data = clean_data)
coef1 <- coef(table3col1)[1:2]
col1 <- c(coef1,count(clean_data))
table3 <- cbind(col1)
colnames(table3) = c("ALL")
rownames(table3) = c("constant(control mean)", "first year GPA < cutoff", "observations")
print(table3)
```

```{r}
colnames(clean_data)
```

```{r}
class(clean_data$hsgrade_pct)
head(clean_data$highHS)
```
```{r}
reg_ex <- felm(gradin4~gpalscutoff+gpaXgpalscutoff+
                 gpaXgpagrcutoff|0|0|clustervar,
               data = subset(raw_data, dist_from_cut >= -0.6 & 
                               dist_from_cut <= 0.6 & male == 1))
```
```{r}
summary(reg_ex)
```