### 2021-04-22 model_test_script.Rmd
Commit for README and Discussion

---
title: "Untitled"
author: "Hong Shi"
date: "4/17/2021"
output: pdf_document
---
```{r}
library(tidyverse)
library(here)
library(finalfit)
library(kableExtra)
library(rdrobust)
library(broom)
#install.packages("rdpower")
library(rdpower)
```

```{r}
clean_cut <- read_csv(here("inputs/data/adjust_bandwidth_data.csv"))
```

```{r}
# Histogram of the distribution of GPA from cutoff

ggplot(clean_for_summary, aes(x = dist_from_cut))+
  geom_histogram(aes(y  = stat(count)), binwidth = 0.125, color = "black", fill = "#33CCCC")+
  geom_vline(xintercept = 0, linetype = "dotted")+
  theme_minimal()+
  theme(plot.title = element_text(hjust = 0.5)) +
  ylab("Frequency Headcount")+
  xlab("Year 1 GPA minus Probation Cutoff")+
  ggtitle("Distribution of Year 1 GPA relative to Probation Cutoff for All Students")
```

```{r}
# plot to show the probation assignment
ggplot(clean_cut,aes(x = dist_from_cut, y = status, group = status, color = status))+
  geom_point(alpha = 0.1)+
  geom_smooth(method = "lm")+
  theme_minimal()+
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5))+
  labs( title = "Stuent Academic Status",
        y ="Academic Status",
        x = "Year 1 GPA minus Probation Cutoff")
```
```{r}
# Check the number of students with administrative data error reporting 
clean_for_summary %>% 
  filter((dist_from_cut < 0 & probation_year1 !=1)| (dist_from_cut >=0 & probation_year1 !=0) ) %>% 
  length()
```

```{r}
clean_cut$status <- as.factor(clean_cut$probation_year1)
ggplot(data= clean_cut, aes(x = dist_from_cut, y = GPA_year2, color = status))+
  geom_point(alpha = 0.1)+
  geom_smooth()
```
```{r}
class(clean_cut$gender)
class(clean_cut$highHS)
class(clean_cut$first_language)
clean_cut$gender <- as.factor(clean_cut$gender)
clean_cut$highHS <- as.factor(clean_cut$highHS)
clean_cut$first_language <- as.factor(clean_cut$first_language)

```
```{r}
class(clean_cut$)
class(clean_cut2$dist_from_cut)

```

```{r}
rdplot(y = clean_cut$GPA_year2, x = clean_cut$dist_from_cut, c = 0, p = 1, covs = cbind(clean_cut$gender,clean_cut$highHS, clean_cut$first_language), title = "GPA by Year 4")
```
```{r}
rdplot(y = clean_cut$left_school, x = clean_cut$dist_from_cut, c = 0, p = 1, covs = cbind(clean_cut$gender,clean_cut$highHS, clean_cut$first_language), title = "GPA by Year 4")
```

```{r}
clean_male <- clean_cut %>% 
  filter(female ==1)
class(clean_male$left_school)
clean_male$left_school <- as.factor(clean_male$left_school)
```

```{r}
test_male <- glm(left_school ~ status + dist_from_cut + highHS + first_language ,family = "binomial", data = clean_male)
summary(test_male)
```
```{r}
rdplot(y = clean_male$left_school, x = clean_male$dist_from_cut, c = 0, p = 1, covs = cbind(clean_male$highHS, clean_male$first_language), title = "GPA by Year 4")
```

```{r}
gpa_camp2 <- lm(GPA_year2 ~ status + dist_from_cut + gender + first_language + highHS*study_campus, data = clean_cut)
summary(gpa_camp2)
```

```{r}
dropout_camp2 <- glm(left_school ~ status + dist_from_cut + gender + highHS*study_campus + first_language ,family = "binomial", data = clean_cut)
summary(dropout_camp2)
```

```{r}
grad_camp2 <- glm(gradin4 ~ status + dist_from_cut + gender + highHS*study_campus + first_language ,family = "binomial", data = clean_cut)
summary(grad_camp2)
```

```{r}
clean_cut1 <- clean_cut %>% 
  filter(dist_from_cut >= -0.1, dist_from_cut <= 0.1)

dropout_01 <- glm(left_school ~ status + dist_from_cut + gender + highHS + first_language ,family = "binomial", data = clean_cut1)
summary(dropout_01)
```

```{r}
gpa_01 <- lm(GPA_year2 ~ status + dist_from_cut + gender + highHS + first_language, data = clean_cut1)
summary(gpa_01)
```

```{r}
grad_01 <- glm(gradin4 ~ status + dist_from_cut + gender + highHS + first_language ,family = "binomial", data = clean_cut1)
summary(grad_01)
```
```{r}
clean_for_summary$hsgrade_pct
```

```{r}
ggplot(clean_for_summary, aes(x = hsgrade_pct))+
  geom_histogram(aes(y  = stat(count)), binwidth = 2.5, color = "black", fill = "#33CCCC")+
  theme_minimal()+
  theme(plot.title = element_text(hjust = 0.5)) +
  ylab("Frequency Headcount")+
  xlab("Year 1 GPA minus Probation Cutoff")+
  ggtitle("Distribution of Year 1 GPA relative to Probation Cutoff for All Students")
```

```{r}
ggplot(clean_cut2, aes(x = hsgrade_pct))+
  geom_histogram(aes(y  = (..count..) ), binwidth = 5, color = "black", fill = "#33CCCC")+
  theme_minimal()+
  facet_wrap(~status)+
  theme(plot.title = element_text(hjust = 0.5)) +
  ylab("Frequency Headcount")+
  xlab("Year 1 GPA minus Probation Cutoff")+
  ggtitle("Distribution of Year 1 GPA relative to Probation Cutoff for All Students")
```

```{r}
ggplot(clean_for_summary, aes(x = hsgrade_pct))+
  geom_histogram(aes(y  = stat(count)), binwidth = 5, color = "black", fill = "#33CCCC")+
  theme_minimal()+
  facet_wrap(~status)+
  theme(plot.title = element_text(hjust = 0.5)) +
  ylab("Frequency Headcount")+
  xlab("Year 1 GPA minus Probation Cutoff")+
  ggtitle("Distribution of Year 1 GPA relative to Probation Cutoff for All Students")
```

```{r}
categorical1 <- clean_cut1 %>% 
  summary_factorlist("status", c("gender", "birthplace", "study_campus", "first_language"),
                     p = FALSE, 
                     digits = c(1,1,3,1),
                     add_dependent_label = TRUE, 
                     dependent_label_prefix = "",
                     add_col_totals = TRUE,
                     include_row_missing_col = FALSE,
                     col_totals_rowname = "",
                     col_totals_prefix = "N(%) = ")
continuous1 <- clean_cut1 %>% 
  summary_factorlist("status", c("hsgrade_pct", "totcredits_year1", "age_at_entry"),
                     p = FALSE,
                     digits = c(2,2,3,1),
                     add_dependent_label = TRUE, 
                     dependent_label_prefix = "",
                     add_col_totals = TRUE,
                     include_row_missing_col = FALSE,
                     col_totals_rowname = "",
                     col_totals_prefix = "N(%) = ")
combine1 <- rbind(categorical1, continuous1)
combine1 <- rbind(combine1, c("","","", ""))
combine1 <- rbind(combine1, c("","","Mean (SD)", "Mean (SD)"))
combine1 <- combine1[-c(12),]
combine1$` ` <- replace(combine1$` `, combine1$` ` == "Mean (SD)", "")
combine1$status <- NULL
combine1 <- combine1[c(1,3,2,5,4,6:11,15,16,12,13,14),] %>% 
  mutate(Description = c("","Gender","","Birth Place", "","","Study Campus","","", "First Language","","","","Highschool Grade Percentile","Credits Attempted in First Year", "Age at Entry"))

kable(combine1[,c(4,1,2,3)],caption = "Summary Statistics of Students Within 0.2 GPA of Probation Cutoff",
      booktabs = TRUE, linesep = "") %>% 
  kable_styling(latex_options = "hold_position")

```

```{r}
gpa_camp2 <- lm(GPA_year2 ~ status + dist_from_cut + gender + first_language + highHS + study_campus, data = clean_cut)
summary(gpa_camp2)
```

```{r}
dropout_camp2 <- glm(left_school ~ status + dist_from_cut + gender + first_language + highHS + study_campus, family = "binomial", data = clean_cut)
summary(dropout_camp2)
```

```{r}
grad_camp2 <- glm(gradin4 ~ status + dist_from_cut + gender + first_language + highHS + study_campus  ,family = "binomial", data = clean_cut)
summary(grad_camp2)

```

```{r}
grad_camp2$coefficients
```
```{r}
clean_cut %>% group_by(study_campus) %>% count()
```

```{r}
sum(is.na(clean_cut$study_campus))
```